---
title: 'Reproducible research : Course Project 1'
author: "Tarang Mehrotra"
date: "June 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Loading and preprocessing the data: We change the date column from Factor to date class.

```{r}
features <- read.csv('./activity.csv', header = T, sep = ',')

str(features)

features$date<-as.Date(features$date,"%m/%d/%Y")

str(features)


```

## Including Plots

For calculating the total steps, date-wise after ignoring missing values.

```{r  echo=TRUE}

features1<-features[which(!(is.na(features$steps))),]

steps_total<-tapply(features1$steps,features1$date,sum)

```

Histogram of the total number of steps taken each day:
```{r fig.height=7,echo=TRUE}
hist(steps_total,main="Histogram of total number of steps per day")
```

To Calculate and report the mean and median of the total number of steps taken per day

Median:
```{r echo=TRUE}
summary(steps_total)[3]
```

Mean:
```{r echo=TRUE}
summary(steps_total)[4]
```

Time series plot(type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days:

```{r fig.height=7, echo=TRUE}
steps_mean_interval<-tapply(features1$steps,features1$interval,mean)

plot(steps_mean_interval, type='l', 
     main="Average number of steps per interval for all days", xlab="Interval", 
     ylab="Average number of steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
which.max(steps_mean_interval)
```

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
is.na(features)

sum(is.na(features))
```

Create a new dataset that is equal to the original dataset but with the missing data filled in with the average value of that interval:

```{r warning=FALSE}

features3<-features

features3$steps[which((is.na(features3$steps)))] <- steps_mean_interval[which((is.na(features3$steps)))%%nrow(steps_mean_interval)]
 
steps_total_after_missing<-tapply(features3$steps,features3$date,sum)

steps_total_after_missing
```

Histogram of the new data with missing values replaced.

```{r}
hist(steps_total_after_missing,main="Histogram of total number of steps per day")
```

Mean and Median of the new data set:

```{r}
summary(steps_total_after_missing)
```

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends:

```{r fig.height=8}
x<-(weekdays(features3$date,abbr = T ))
x<-gsub("S(at|un)","Weekend",x)
x<-gsub("Mon|Tue|Wed|Thu|Fri","Weekday",x)

features3$day<-as.factor(x)

avg_step_imp <- aggregate(steps ~ interval + day, data = features3, mean)

xyplot(steps~interval|day,data=avg_step_imp,layout(1,2),type = "l")
```
